Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-25 paper.tex

In this section, we review lazy cast calculi
(Section~\ref{sec:cast-calculi}) and coercions 
(Section~\ref{sec:coercion-calculus}).

@@

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or continuing $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the natural way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

\begin{figure}
  \[
  \begin{array}{lrcl}
  \stxrule{I}{Injectable types (\lazyD)}{
    \POOfun{T}{T} \mid \POOunit}
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOfun{\TOOdyn}{\TOOdyn} \mid \POOunit
  }
  \stxrule{c}{Coercions}{
    \ncInj{I} \mid
    \ncProj{I}{l} \mid
    \ncId \mid
    \ncFail{l} \mid
    \ncSeq{c}{c} \mid
    \ncFun{c}{c}
  }
\stxrule{\hat{c}}{normal coercions}{
  \ncFun{\hat{c}}{\hat{c}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
  \ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
  \ncId \mid
  \ncProj{I}{l} \mid
  \ncInj{I} \mid
  \ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
  \ncFail{l} 
}
  \end{array}
  \]
  \caption{Syntax of coercions and normal forms}
  \label{fig:normal-coercion}
\end{figure}

@@

\figref{fig:HC-D} defines the functions $id$, $seq$, and $cast$ that
construct \lazyD{} hypercoercions, towards satisifying the Cast abstract data 
type defined in Section~\ref{sec:framework}.
\figref{fig:HC-D} also defines some auxiliary functions.

The $seq$ operator is defined in terms of coercion composition, which
is the key to compressing coercions and maintaining space-efficiency.
A composition operator for coercions typically requires the target
type of the first coercion to match the source type of the second.
However, it is useful to relax this restriction for the D blame
tracking strategy.
%
We shall need optional blame labels, written $\ell$ , that range over
$\epsilon$ or $l$. The label is mandatory when the target
type of the first coercion does not match the source type of the second. 
Then we write composition of hypercoercions as $c_1
\fatsemi^\ell c_2$.  When both $c_1$ and $c_2$ are $\hyperCoercionI$,
their composition is also $\hyperCoercionI$.  If the head of $c_2$ is
a projection, the result is $c_2$.  Otherwise, we need a label to
build the projection. Since the head of $c_2$ is the no-op, its
source type must be a pretype. Thus we know $\ell$ must be a label and
we put it in the projection. When $c_1$ ends with a failure, the
composition is $c_1$ itself. In all the remaining cases, $c_1$ does
not end with a failure.  When $m_1$ and $m_2$ have different top
constructors ($m_1 \not\smile m_2$), the result is a failure coercion,
which needs a label. When $c_2$ starts with a projection, we blame the 
projection for casting a value to a shallowly inconsistent type. When $c_2$
starts with the no-op, we know $\ell$ must be a label, which is blamed for
composing shallowly inconsistent hypercoercions. The last two cases
compose $m_1$ and $m_2$ with the auxilliary function $m \fatsemi^\ell
m$, which assumes its inputs have the same top constructor. The
definition of $m \fatsemi^\ell m$ is a straightforward structural
recursion. Going back to the last two cases of $c_1 \fatsemi^\ell
c_2$. When $c_2$ starts with a projection, we do not know whether the
target type of $m_1$ matches the source of $m_2$. So we use the $l$ in the 
projection to  compose the middles $m_1 \fatsemi^l m_2$. When $c_2$ starts with 
the no-op, we compose the middles using $\ell$.

@@

Hypercoercions for the UD blame tracking strategy have the same syntax
as for D (\figref{fig:hypercoercion}), but the definitions of $id$,
$cast$, and $seq$, differ from D.  Again, the $seq$ operator
(Figure~\ref{fig:HC-UD}) is defined in terms of composition, but here
composition $c_1 \fatsemi c_2$ makes the usual assumption that the
target type of $c_1$ matches the source type of $c_2$.  The definition
of composition for UD is particularly straightfoward. The first two
lines say that $\hyperCoercionI$ acts as the identity on both the left
and right. The third line handles the case when both the tail of $c_1$
and the head of $c_2$ are no-ops, in which case the middles types
$m_1$ and $m_2$ are composed via an auxiliary composition
operator. This compose operator for middle coercions assumes that its
inputs are shallowly consistent.  The fourth line handles the
important case when the tail of $c_1$ is an injection and the head of
$c_2$ is a projection. If the two middle coercions are shallowly
consistent, then they can be composed. If not, the tail of the result
is a failure coercion with the projection's label.  The last line
handles the case when the tail of $c_1$ is a failure, in which case
the result of composition is $c_1$.

@@

Hypercoercions enable a bit-level representation that is particularly
compact for the most common-case coercions. The most frequent
coercions are injections and projections to base types (here
$\POOunit$ is the only base type, but in a real language the base
types would include integers, Booleans, etc.) and identity coercions.
Values of base types are often stored in CPU registers, so it would be
nice for coercions on base types to also fit in registers, that is, in
64 bits, so that applying a coercion to a value would not require
access to main memory, which is an order of magnitude slower than
accessing registers on a modern CPU.
%
Coercions involving non-base types, such as functions types,
may be arbitrarily large, so in those cases the hypercoercion
representation can be a pointer to a heap-allocated structure.  Here
is a sketch for the bit-level representation for hypercoercions.

@@

\item If the middle is of base type, then the remaining 62 bits are
  used to represent the head (25 bits), middle (11 bits), and tail (26
  bits).
  \begin{itemize}
  \item The head requires 1 bit to differentiate between $\epsilon$
    and $?^l$ and then 24 bits could be used for the label $l$.
  \item The middle would use 11 bits to differentiate all the base types.
  \item The tail requires 2 bits to differentiate between $\epsilon$,
    $!$, and $\bot^l$, and could use 24 bits for the label $l$.
  \end{itemize}
  In the event that a program requires too many blame labels, then one
  can fallback to using the heap-allocated structure for more
  coercions.
\end{itemize}

@@

The framework includes an abstract data type named Cast ADT 
(Section~\ref{sec:framework:adt}), and a CEK machine \effCEK{C} that is 
parameterized over Cast ADTs (Section~\ref{sec:framework:cek}). This machine is 
space-efficient provided that the Cast ADT implementation performs compression. 
We conjecture that all cast representations defined in the literature
are instances of this ADT.

@@

In Section~\ref{secc:framework:monoid-correct} we prove that, for any
instance $C$ of the Cast ADT, if $C$ satisfies a more refined abstract
data type for Lazy D casts, then \effCEK{C} is equivelent to
\ineffCEKD{}, that is,
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]
We conjecture that all of the cast representations in the literature
for \lazyD{} (supercoercions, coercions in normal form, threesomes)
are instances of the Lazy D Cast ADT.  We are working on a similar
theorem for \lazyUD{} cast representations.

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all casted values.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push casts into the ordinary 
values. Thus non-casted values in \ineffCEK\ corresponds to values in 
\effCEK{C} where casts are identities. For instance, $(\vOOcons{v_1}{v_2})$ 
corresponds to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$.

@@

The \emph{Cast Abstract Data Type}, defined below, captures the set of
operators that a cast representation $C$ must provide for it to be
used with the \effCEK{C} machine.  The first three operators enable
\effCEK{C} to construct casts so we call them \textit{cast
  constructors}. The forth and last operator enables \effCEK{C} to
apply casts to values.

@@

Next we define the \emph{Lazy D Cast ADT}, which capture the further
requirements that are needed for the theorem that establishes
equivalence to \ineffCEKD{}. Property (1) below states that $id(T)$
acts like the identity function. Property (2) states that $seq(c,c)$
acts like a sequence of casts. Properties (3) through (11) state that
$cast(T_1,l,T_2)$ act like the Lazy D $applyCast$ for \ineffCEK\
(Fig.~\ref{fig:applyCast-D-C}).

@@

\figref{fig:machine-cekcc} defines \effCEK{C}, a space-efficient CEK
machine parameterized over Cast ADT ($C$).  The key differences between 
\effCEK{C} and \ineffCEK\ are the places where casts can accumulate at runtime, 
that is, values and continuations. We discuss above the difference in values
(Definition~\ref{def:values-effCEK}).

\begin{definition}[Continuations for the \effCEK{C} machine] 
\label{def:values-effCEK}
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]
\end{definition}

@@

The transition relation $\judgeSreduce{C}{s}{s}$ relies on functions provided 
by $C$ to work with casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts with $id$s.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$cont$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluate the function body.
%
To take out the first part of a pair, the machine returns the first value to a 
continuation extended with the first cast. Taking out the second 
part of a pair is similar.
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
To case split a variant, the machine first look at the variant value. If the 
variant is a left injection, the machine composes the cast in the left 
injection with the domain cast of the first continuation function, then move 
to a state that will apply the new function to the value in the left injection.
The case for right injections is similar.

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both side.
%
Rules for products and sums are similar.

@@

Other cases are similar. Of course we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case spliting, we employ $cast(T,l,T)=id(T)$ to relate the continuation 
functions.
\end{proof}

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:SD-equiv}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid,
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
	By the definitions of $eval$, we need to show 
	\[
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}}{
		\sOOhalt{o}
	}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{\kOOmt}}{
		\sOOhalt{o}
	}
	\]
	Obviously, the initial states are related in the bisimulation relation.
	\[
	\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}
	\approx
	\sOOinspect{e}{\emptyset}{\kOOmt}
	\]
	So we can generalize our goal, and prove that if $s_1 \approx s_2$
	\[
	\judgeSreduceTrans{C}{s_1}{\sOOhalt{o}}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{s_2}{\sOOhalt{o}}
	\]

@@

\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
  \label{thm:CEKS-bisim}
  Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
  $s_{1} \in \text{\effCEK{C_1}}$,
  $s_{2} \in \text{\effCEK{C_2}}$,
  $s_{1} \approx s_{2}$,
  \begin{enumerate}
    \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
    \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
    \item If $\judgeSreduce{C_1}{s_1}{s_3}$ 
    then $\judgeSreduce{C_2}{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
    \item If $\judgeSreduce{C_2}{s_2}{s_4}$ 
    then $\judgeSreduce{C_1}{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
  \end{enumerate}
% 
  Or diagrammatically, 
  \[\begin{array}{clc}
  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_3\\
  \rotatebox[origin=c]{90}{$\approx$} 
  & & \rotatebox[origin=c]{90}{$\approx$} \\
  s_{2} & 
  \longrightarrow_{\text{\ineffCEKD}} & s_4 \\
  \end{array}\]
\end{lemma}
\begin{proof} This proof is straightforward. The key ideas are undoing 
  uses of $seq$ with the property (2) of Lazy D Cast ADT, and handling all
  uses of $cast$ with properties (3)-(11).
\end{proof}

>>>>

In this section, we fist review lazy cast calculi
(Section~\ref{sec:cast-calculi}), where we present an abstract machine that is 
employed in our framework (Section~\ref{sec:framework}). Then we review 
coercions and their normal forms (Section~\ref{sec:coercion-calculus}), which 
motivates and inspires the design of hypercoercions 
(Section~\ref{sec:hypercoercion-definition}).

@@

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values $v$ and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machines involve two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an evaluating
$\sOOinspect{e}{\mathcal{E}}{\kappa}$ or continuing $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation ($o$). Observations include all value 
constructors and blame. The function converting values to 
observations ($observe(v) = o$) is defined in the obvious way.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

@@

\begin{figure}
  \[
  \begin{array}{lrcl}
  \stxrule{I}{Injectable types (\lazyD)}{
    \POOfun{T}{T} \mid \POOunit}
  \stxrule{I}{Injectable types (\lazyUD)}{
    \POOfun{\TOOdyn}{\TOOdyn} \mid \POOunit
  }
  \stxrule{c}{Coercions}{
    \ncInj{I} \mid
    \ncProj{I}{l} \mid
    \ncId \mid
    \ncFail{l} \mid
    \ncSeq{c}{c} \mid
    \ncFun{c}{c}
  }
\stxrule{\hat{c}}{normal coercions}{
  \ncFun{\hat{c}}{\hat{c}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncFun{\hat{c}}{\hat{c}}} \mid
  \ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncSeq{\ncFun{\hat{c}}{\hat{c}}}{\ncInj{I}}} \mid
  \ncSeq{\ncProj{I}{l}}{\ncFail{l}}
}
\stxrulecont{
  \ncId \mid
  \ncProj{I}{l} \mid
  \ncInj{I} \mid
  \ncSeq{\ncProj{I}{l}}{\ncInj{I}} \mid
  \ncFail{l} 
}
  \end{array}
  \]
  \caption{Syntax of coercions and normal forms {\`a} la \citet{siek2012interpretations}.}
  \label{fig:normal-coercion}
\end{figure}

@@

\figref{fig:HC-D} defines the functions $id$, $seq$, and $cast$ that
construct \lazyD{} hypercoercions, towards satisfying the Cast abstract data 
type defined in Section~\ref{sec:framework}.
\figref{fig:HC-D} also defines some auxiliary functions.

The $seq$ operator is defined in terms of coercion composition, which
is the key to compressing coercions and maintaining space-efficiency.
A composition operator for coercions typically requires the target
type of the first coercion to match the source type of the second.
However, it is useful to relax this restriction for the D blame
tracking strategy.
%
We shall need optional blame labels, written $\ell$ , that range over
$\epsilon$ or $l$. The label is mandatory when the target
type of the first coercion does not match the source type of the second. 
Then we write the composition of hypercoercions as $c_1
\fatsemi^\ell c_2$.  When both $c_1$ and $c_2$ are $\hyperCoercionI$,
their composition is also $\hyperCoercionI$.  If the head of $c_2$ is
a projection, the result is $c_2$.  Otherwise, we need a label to
build the projection. Since the head of $c_2$ is the no-op, its
source type must be a pretype. Thus we know $\ell$ must be a label and
we put it in the projection. When $c_1$ ends with a failure, the
composition is $c_1$ itself. In all the remaining cases, $c_1$ does
not end with a failure.  When $m_1$ and $m_2$ have different top
constructors ($m_1 \not\smile m_2$), the result is a failure coercion,
which needs a label. When $c_2$ starts with a projection, we blame the 
projection for casting a value to a shallowly inconsistent type. When $c_2$
starts with the no-op, we know $\ell$ must be a label, which is blamed for
composing shallowly inconsistent hypercoercions. The last two cases
compose $m_1$ and $m_2$ with the auxiliary function $m \fatsemi^\ell
m$, which assumes its inputs have the same top constructor. The
definition of $m \fatsemi^\ell m$ is a straightforward structural
recursion. Going back to the last two cases of $c_1 \fatsemi^\ell
c_2$. When $c_2$ starts with a projection, we do not know whether the
target type of $m_1$ matches the source of $m_2$. So we use the $l$ in the 
projection to  compose the middles $m_1 \fatsemi^l m_2$. When $c_2$ starts with 
the no-op, we compose the middles using $\ell$.

@@

Hypercoercions for the UD blame tracking strategy have the same syntax
as for D (\figref{fig:hypercoercion}), but the definitions of $id$,
$cast$, and $seq$, differ from D.  Again, the $seq$ operator
(Figure~\ref{fig:HC-UD}) is defined in terms of composition, but here
composition $c_1 \fatsemi c_2$ makes the usual assumption that the
target type of $c_1$ matches the source type of $c_2$.  The definition
of composition for UD is particularly straightforward. The first two
lines say that $\hyperCoercionI$ acts as the identity on both the left
and right. The third line handles the case when both the tail of $c_1$
and the head of $c_2$ are no-ops, in which case the middles types
$m_1$ and $m_2$ are composed via an auxiliary composition
operator. This compose operator for middle coercions assumes that its
inputs are shallowly consistent.  The fourth line handles the
important case when the tail of $c_1$ is an injection and the head of
$c_2$ is a projection. If the two middle coercions are shallowly
consistent, then they can be composed. If not, the tail of the result
is a failure coercion with the projection's label.  The last line
handles the case when the tail of $c_1$ is a failure, in which case
the result of composition is $c_1$.

@@

Hypercoercions enable a bit-level representation that is particularly
compact for the most common-case coercions. The most frequent
coercions are injections and projections to base types (here
$\POOunit$ is the only base type, but in a real language the base
types would include integers, Booleans, etc.) and identity coercions.
Values of base types are often stored in CPU registers, so it would be
nice for coercions on base types to also fit in registers, that is, in
64 bits, so that applying a coercion to a value would not require
access to main memory, which is an order of magnitude slower than
accessing registers on a modern CPU.
%
Coercions involving non-base types, such as function types,
may be arbitrarily large, so in those cases, the hypercoercion
representation can be a pointer to a heap-allocated structure.  Here
is a sketch for the bit-level representation for hypercoercions.

@@

\item If the middle is of base types, then the remaining 62 bits are
  used to represent the head (25 bits), middle (11 bits), and tail (26
  bits).
  \begin{itemize}
  \item The head requires 1 bit to differentiate between $\epsilon$
    and $?^l$ and then 24 bits could be used for the label $l$.
  \item The middle would use 11 bits to differentiate all the base types.
  \item The tail requires 2 bits to differentiate between $\epsilon$,
    $!$, and $\bot^l$, and could use 24 bits for the label $l$.
  \end{itemize}
  In the event that a program requires too many blame labels, then one
  can fallback to using the heap-allocated structure for more
  coercions.
\end{itemize}

@@

The framework includes an abstract data type named Cast ADT 
(Section~\ref{sec:framework:adt}), and a CEK machine \effCEK{C} that is 
parameterized over Cast ADTs (Section~\ref{sec:framework:cek}). This machine is 
space-efficient provided that the Cast ADT implementation performs compression. 
We conjecture that all cast representations defined in the literature
are instances of this ADT. The framework is available at the following URL:
\[\text{\urlLazyDProof}\]

@@

In Section~\ref{secc:framework:monoid-correct} we prove that, for any
instance $C$ of the Cast ADT, if $C$ satisfies a more refined abstract
data type for Lazy D casts, then \effCEK{C} is equivalent to
\ineffCEKD{}, that is,
\[
  \evalEqv{\effCEK{C}}{\ineffCEKD}
\]
We conjecture that all of the cast representations in the literature
for \lazyD{} (supercoercions, coercions in normal form, threesomes)
are instances of the Lazy D Cast ADT.  We are working on a similar
theorem for \lazyUD{} cast representations.

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all casted values.  In \effCEK{C}, however, we do not
have this generic value constructor, instead, we push casts into the ordinary 
values. Thus non-casted values in \ineffCEK\ correspond to values in 
\effCEK{C} where casts are identities. For instance, $(\vOOcons{v_1}{v_2})$ 
corresponds to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$.

@@

The \emph{Cast Abstract Data Type}, defined below, captures the set of
operators that a cast representation $C$ must provide for it to be
used with the \effCEK{C} machine.  The first three operators enable
\effCEK{C} to construct casts so we call them \textit{cast
  constructors}. The fourth and last operator enables \effCEK{C} to
apply casts to values.

@@

Next, we define the \emph{Lazy D Cast ADT}, which captures the further
requirements that are needed for the theorem that establishes
equivalence to \ineffCEKD{}. Property (1) below states that $id(T)$
acts as the identity function. Property (2) states that $seq(c,c)$
acts as a sequence of casts. Properties (3) through (11) state that
$cast(T_1,l,T_2)$ acts like the Lazy D $applyCast$ for \ineffCEK\
(Fig.~\ref{fig:applyCast-D-C}).

@@

%<<<<<<< HEAD
\effCEK{C} is a space-efficient CEK machine parameterized over Cast ADT ($C$).  
The key differences between \effCEK{C} and \ineffCEK\ are the places where 
casts can accumulate at runtime, that is, values and continuations. We discuss 
above the difference in values (Definition~\ref{def:values-effCEK}).

\begin{definition}[Continuations for the \effCEK{C} machine] 
\label{def:values-effCEK}
%=======
%\begin{figure}
%%  Syntax
%>>>>>>> bd39888b0287dbef61e981bd0e79bc8822fc85ce
  \[
  \begin{array}{lrcl}
% \stxrule{r}{cast results}{
%   \rOOsucc{v} \mid
%   \rOOfail{l}
% }
% \stxrule{s}{states}{
%   \sOOinspect{e}{\mathcal{E}}{\kappa} \mid{}
%   \sOOreturn{v}{\kappa} \mid{}
%   \sOOhalt{o}
% }
  \stxrule{\kappa}{Continuation}{
    \kOOcast{c}{k}
  }
  \stxrule{k}{Pre-continuations}{
    \kOOmt \mid{}
    \kOOconsI{T}{T}{e}{\mathcal{E}}{\kappa} \mid
    \kOOconsII{T}{T}{v}{\kappa} \mid
    \kOOinl{T}{T}{\kappa}
  }
  \stxrulecont{
  	\kOOinr{T}{T}{\kappa} \mid
    \kOOappI{e}{\mathcal{E}}{\kappa} \mid
    \kOOappII{v}{\kappa} \mid
    \kOOcar{\kappa} \mid
    \kOOcdr{\kappa}
  }
  \stxrulecont{
    \kOOcaseI{e}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseII{v}{e}{\mathcal{E}}{\kappa} \mid
    \kOOcaseIII{v}{v}{\kappa}
  }
  \end{array}
  \]
\end{definition}

@@

\figref{fig:machine-cekcc} defines the transition relation 
$\judgeSreduce{C}{s}{s}$. They rely on functions provided 
by $C$ to work with casts.
%
When evaluating a cast expression, the machine extends the continuation with 
cast $cast(T_1,l,T_2)$. The function $ext(c,\kappa)$ composes $c$ with the cast 
at the top of $k$ by calling $seq$.
%
To construct a function, the machine fills the casts with $id$s.
%
To return a value to a continuation, the machine first applies the top cast to 
the value. If the application fails, the machine halts with the blame label 
from the failure. Otherwise, the machine handles the pre-continuation with 
$cont$.
%
When the machine constructs a pair, a left injection, or a right injection, it 
fills the casts with $id$ as well, just like how it did for functions. 
%
To apply a function, the machine applies the domain cast $c_1$ to the 
operand $v_2$, extends the continuation with the codomain cast $c_2$, and 
evaluates the function body.
%
To take out the first part of a pair, the machine returns the first value to a 
continuation extended with the first cast. Taking out the second 
part of a pair is similar.
%
%%In case splitting, if the target value is a left injection, the
%%machine moves to a state that will apply the first continuation
%%function to the value inside the left injection. The case for right
%%injection is similar.
%%
To case split a variant, the machine first looks at the variant value. If the 
variant is a left injection, the machine composes the cast in the left 
injection with the domain cast of the first continuation function, then move 
to a state that will apply the new function to the value in the left injection.
The case for right injections is similar.

@@

Fig.~\ref{fig:bisim-SC-D} defines the bisimulation relation.  It is
the smallest congruence relation that also includes the pairs induced
by the rules in Fig.~\ref{fig:bisim-SC-D}.
%
The relation is mutually defined on continuations, written $\kappa
\approx \kappa : T$, where $T$ is their source type, and values, written $v 
\approx v : T$, where $T$ is their type.
%
If $k \approx \kappa : T$, adding an identity cast on top of $k$ gives a 
related continuation.
%
%\todo[inline]{The use of ``wrap'' obscures this explanation.
%  Consider getting rid of wrap by inlining it. -JS
%\\---\\
%I agree that inlining wrap and ext can improve readability of $\kappa \approx 
%\kappa$, but this would make latex different from Agda, is it alright?
%-KC
%\\---\\
%Yes, that is alright. Inlining is really a trivial change.
%-JS
%}
%
If $\kOOcast{c}{k_1} \approx \kappa_2 : T$, extending them with 
the same cast gives related continuations.
%
If two values of type $P$ are related, injecting them to $\TOOdyn$ gives 
related values.
%
% The following is part of being a congruence. -JS
%Unit values are of course related. 
%
To relate to a naked function in \ineffCEKD, the function in \effCEK{C} 
must use identities as its casts.
%
If functions are related, we can apply the same cast on both sides.
%
Rules for products and sums are similar.

@@

Other cases are similar. Of course, we need the lemma about $applyCast$ 
(\ref{thm:lem-apply-cast-SD}) when cast applications happen in both machines. 
And in case splitting, we employ $cast(T,l,T)=id(T)$ to relate the continuation 
functions.
\end{proof}

\begin{corollary}[Correctness of \effCEK{C}]
  \label{thm:SD-equiv}
  Suppose $C$ is an instance of the Cast ADT
  and the \lazyD{} ADT, $C$ is a monoid
  and $cast(T,l,T)=id(T)$.
  If $\judgetype{\emptyset}{e}{T}$ and $o : T$, then
  \[
    \text{\evalEqv{\effCEK{C}}{\ineffCEKD}}
  \]
\end{corollary}
\begin{proof}
	By the definitions of $eval$, we need to show 
	\[
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}}{
		\sOOhalt{o}
	}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{\sOOinspect{e}{\emptyset}{\kOOmt}}{
		\sOOhalt{o}
	}
	\]
	Obviously, the initial states are related in the bisimulation relation.
	\[
	\sOOinspect{e}{\emptyset}{wrap(\kOOmt)}
	\approx
	\sOOinspect{e}{\emptyset}{\kOOmt}
	\]
	So we can generalize our goal, and prove that if $s_1 \approx s_2$
	\[
	\judgeSreduceTrans{C}{s_1}{\sOOhalt{o}}
	\;\text{if and only if}\;
	\judgeSreduceTrans{C}{s_2}{\sOOhalt{o}}
	\]

@@

\begin{lemma}[Strong Bisimulation among $\mathcal{S}(\cdot)$]
  \label{thm:CEKS-bisim}
  Assume $C_1$ and $C_2$ implements Lazy D Cast ADT
  $s_{1} \in \text{\effCEK{C_1}}$,
  $s_{2} \in \text{\effCEK{C_2}}$,
  $s_{1} \approx s_{2}$,
  \begin{enumerate}
    \item If $s_1 = \sOOhalt{o}$ then $s_2 = \sOOhalt{o}$
    \item If $s_2 = \sOOhalt{o}$ then $s_1 = \sOOhalt{o}$
    \item If $\judgeSreduce{C_1}{s_1}{s_3}$ 
    then $\judgeSreduce{C_2}{s_2}{s_4}$ and $s_3 \approx s_4$ for some $s_4$
    \item If $\judgeSreduce{C_2}{s_2}{s_4}$ 
    then $\judgeSreduce{C_1}{s_1}{s_3}$ and $s_3 \approx s_4$ for some $s_3$
  \end{enumerate}
% 
%  Or diagrammatically, 
%  \[\begin{array}{clc}
%  s_{1} & \longrightarrow_{\text{\effCEK{C}}} & s_3\\
%  \rotatebox[origin=c]{90}{$\approx$} 
%  & & \rotatebox[origin=c]{90}{$\approx$} \\
%  s_{2} & 
%  \longrightarrow_{\text{\ineffCEKD}} & s_4 \\
%  \end{array}\]
\end{lemma}
\begin{proof} This proof is straightforward. The key ideas are undoing 
  uses of $seq$ with the property (2) of Lazy D Cast ADT, and handling all
  uses of $cast$ with properties (3)-(11).
\end{proof}
